# Statistical Analysis Plan — AgriStory Evaluation

## R Markdown Template

This document provides reproducible R code for the primary and secondary analyses.

```r
---
title: "AgriStory Evaluation: Statistical Analysis"
author: "Research Team"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: show
    toc: true
    toc_depth: 3
params:
  data_path: "data/study_data.csv"
  output_path: "results/"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, fig.width = 8, fig.height = 6)
library(tidyverse)
library(lme4)
library(emmeans)
library(pwr)
library(ggplot2)
library(knitr)
library(kableExtra)
set.seed(42)
```

## 1. Load Data

```{r load-data}
data <- read_csv(params$data_path)

# Data cleaning
data <- data %>%
  mutate(
    condition = factor(condition, levels = c("conventional", "voice-only", "voice-icons", "voice-story")),
    literacy_group = factor(literacy_group, levels = c("low", "semi", "extension")),
    task_id = factor(task_id),
    user_id = factor(user_id),
    completed = as.logical(completed)
  )

glimpse(data)
```

## 2. Descriptive Statistics

```{r descriptive}
# Completion rate by condition
completion_by_condition <- data %>%
  group_by(condition) %>%
  summarise(
    n = n(),
    completed = sum(completed),
    completion_rate = mean(completed) * 100,
    mean_time = mean(duration_ms, na.rm = TRUE) / 1000,
    sd_time = sd(duration_ms, na.rm = TRUE) / 1000
  )

kable(completion_by_condition, digits = 2, caption = "Completion Rate by Condition")
```

## 3. Primary Analysis: H1 (Comprehension Improvement)

### 3.1 Independent Samples t-test (A vs. D)

```{r h1-ttest}
# Filter for low-literacy participants only (primary analysis)
primary_data <- data %>% filter(literacy_group == "low")

# Comprehension score by condition
comp_scores <- primary_data %>%
  group_by(user_id, condition) %>%
  summarise(comprehension_score = mean(quiz_score, na.rm = TRUE), .groups = "drop")

# T-test: Conventional (A) vs. Voice+Icon+Story (D)
a_vs_d <- comp_scores %>% filter(condition %in% c("conventional", "voice-story"))

t_result <- t.test(comprehension_score ~ condition, data = a_vs_d, var.equal = FALSE)
print(t_result)

# Effect size (Cohen's d)
cohens_d <- function(x, y) {
  nx <- length(x)
  ny <- length(y)
  sx <- sd(x)
  sy <- sd(y)
  pooled_sd <- sqrt(((nx - 1) * sx^2 + (ny - 1) * sy^2) / (nx + ny - 2))
  (mean(x) - mean(y)) / pooled_sd
}

effect_size <- cohens_d(
  a_vs_d$comprehension_score[a_vs_d$condition == "voice-story"],
  a_vs_d$comprehension_score[a_vs_d$condition == "conventional"]
)
cat("Cohen's d:", round(effect_size, 3))
```

### 3.2 ANCOVA (Controlling for Pre-Literacy)

```{r h1-ancova}
# Merge with pre-literacy scores
pre_scores <- data %>% distinct(user_id, pre_literacy_score)
ancova_data <- comp_scores %>% left_join(pre_scores, by = "user_id")

ancova_model <- aov(comprehension_score ~ condition + pre_literacy_score, data = ancova_data)
summary(ancova_model)

# Post-hoc comparisons
emmeans(ancova_model, pairwise ~ condition, adjust = "bonferroni")
```

## 4. Primary Analysis: H2 (Trust/SUS Improvement)

```{r h2-trust}
trust_data <- data %>%
  distinct(user_id, condition, .keep_all = TRUE) %>%
  filter(!is.na(trust_score), !is.na(sus_score))

# Trust score t-test
trust_t <- t.test(trust_score ~ condition, data = trust_data %>% filter(condition %in% c("conventional", "voice-story")))
print(trust_t)

# SUS score t-test
sus_t <- t.test(sus_score ~ condition, data = trust_data %>% filter(condition %in% c("conventional", "voice-story")))
print(sus_t)
```

## 5. Primary Analysis: H3 (Decision Accuracy)

```{r h3-accuracy}
# Decision accuracy by condition
accuracy_data <- data %>%
  group_by(condition) %>%
  summarise(
    accuracy = mean(decision_correct, na.rm = TRUE) * 100,
    n = n()
  )

kable(accuracy_data, digits = 2, caption = "Decision Accuracy by Condition")

# Chi-square test
chi_sq <- chisq.test(table(data$condition, data$decision_correct))
print(chi_sq)
```

## 6. Primary Analysis: H4 (Literacy Moderation)

```{r h4-moderation}
# Moderation analysis: Condition × Literacy Group
mod_data <- comp_scores %>%
  left_join(pre_scores, by = "user_id") %>%
  left_join(data %>% distinct(user_id, literacy_group), by = "user_id")

# Two-way ANOVA
mod_anova <- aov(comprehension_score ~ condition * literacy_group + pre_literacy_score, data = mod_data)
summary(mod_anova)

# Plot interaction
interaction_plot <- ggplot(mod_data, aes(x = condition, y = comprehension_score, color = literacy_group, group = literacy_group)) +
  geom_line(stat = "summary", fun = mean, linewidth = 1.2) +
  geom_point(stat = "summary", fun = mean, size = 3) +
  labs(title = "Comprehension Score by Condition and Literacy Group",
       x = "Condition", y = "Mean Comprehension Score") +
  theme_minimal()

print(interaction_plot)
```

## 7. Secondary Analysis: Four Conditions ANOVA

```{r secondary-anova}
# One-way ANOVA for all 4 conditions
anova_result <- aov(comprehension_score ~ condition, data = comp_scores)
summary(anova_result)

# Post-hoc Tukey HSD
tukey_result <- TukeyHSD(anova_result)
print(tukey_result)

# Visualization
ggplot(comp_scores, aes(x = condition, y = comprehension_score, fill = condition)) +
  geom_boxplot() +
  labs(title = "Comprehension Score Distribution by Condition",
       x = "Condition", y = "Comprehension Score") +
  theme_minimal() +
  theme(legend.position = "none")
```

## 8. Secondary Analysis: Task Difficulty

```{r task-difficulty}
task_difficulty <- data %>%
  group_by(task_id, condition) %>%
  summarise(
    completion_rate = mean(completed, na.rm = TRUE) * 100,
    mean_time = mean(duration_ms, na.rm = TRUE) / 1000
  ) %>%
  ungroup()

ggplot(task_difficulty, aes(x = task_id, y = completion_rate, fill = condition)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Completion Rate by Task and Condition",
       x = "Task", y = "Completion Rate (%)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## 9. Secondary Analysis: Voice Recognition Accuracy

```{r voice-accuracy}
# Calculate WER (Word Error Rate) for transcribed voice inputs
voice_data <- data %>% filter(!is.na(wer))

wer_summary <- voice_data %>%
  group_by(condition, language) %>%
  summarise(
    mean_wer = mean(wer, na.rm = TRUE),
    median_wer = median(wer, na.rm = TRUE),
    n = n()
  )

kable(wer_summary, digits = 3, caption = "Word Error Rate by Condition and Language")

# Correlation: WER vs. comprehension
cor_test <- cor.test(voice_data$wer, voice_data$comprehension_score, use = "complete.obs")
print(cor_test)
```

## 10. Secondary Analysis: Time to Answer

```{r time-analysis}
# Survival analysis for task completion time
library(survival)

# Create survival object (time to completion)
surv_data <- data %>%
  mutate(time_seconds = duration_ms / 1000) %>%
  filter(!is.na(time_seconds))

# Kaplan-Meier curves by condition
km_fit <- survfit(Surv(time_seconds, completed) ~ condition, data = surv_data)
plot(km_fit, col = 1:4, lty = 1:4,
     main = "Task Completion Time by Condition",
     xlab = "Time (seconds)", ylab = "Probability of Completion")
legend("bottomright", legend = levels(factor(surv_data$condition)), col = 1:4, lty = 1:4)
```

## 11. Assumption Checks

```{r assumptions}
# Normality test (Shapiro-Wilk)
shapiro.test(residuals(ancova_model))

# Homogeneity of variance (Levene's test)
library(car)
leveneTest(comprehension_score ~ condition, data = comp_scores)

# Outlier detection
outliers <- boxplot.stats(comp_scores$comprehension_score)$out
cat("Outliers detected:", length(outliers), "\n")
```

## 12. Missing Data Analysis

```{r missing-data}
# Missingness pattern
missing_summary <- data %>%
  select(comprehension_score, trust_score, sus_score, duration_ms) %>%
  summarise(across(everything(), ~ sum(is.na(.))))

kable(missing_summary, caption = "Missing Data Summary")

# Exclude participants with >20% missing tasks
```

## 13. Sensitivity Analysis

```{r sensitivity}
# Compare complete cases vs. imputed
complete_cases <- data %>% drop_na(comprehension_score, trust_score, sus_score)

# Report both complete-case and ITT (intention-to-treat) estimates
```

## 14. Reproducibility

```{r reproducibility}
sessionInfo()
```

---

## Analysis Notes

- All tests are two-tailed with α = 0.05.
- Bonferroni correction applied for multiple comparisons in post-hoc tests.
- Effect sizes reported: Cohen's d for t-tests, partial η² for ANOVA.
- Non-parametric alternatives used if normality assumptions violated.
- All raw data and analysis code will be made available upon publication.
